R for Clinical Study Reports and Submission

### 2021-08-29 tlf-specific-ae.Rmd
place holder for ae tables
add specific AE example

# AE Summary

Following [ICH E3 guidance](https://database.ich.org/sites/default/files/E3_Guideline.pdf),
we need to summarize which patients were included in each efficacy analysis 
in the section 11.1, data sets analysed of a CSR.

@@

>>>>

# Specific AE

Following [ICH E3 guidance](https://database.ich.org/sites/default/files/E3_Guideline.pdf),
we need to summarize which patients were included in each efficacy analysis 
in the section 12.2, adverse events of a CSR.

@@

In this chapter, we illustrate how to summarize simplified specific AE information in a study.

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("tlf/tlf_spec_ae.pdf")
```

The data used to summarize AE information is in `adsl` and `adae` datasets.

```{r}
adsl <- read_sas("adam_data/adsl.sas7bdat")
adae <- read_sas("adam_data/adae.sas7bdat")
```

For illustration purpose, we only provide count in the simplified table. 
The percentage of participants 
for each AE criteria can be calculated as in Chapter \@ref(aesummary).

In this way, let's focus on the analysis script for advanced feature for table layout.

- group content: AE can be summarized in multiple nested layer. 
(e.g. by system organ class (SOC, `AESOC`) and specific AE term (`AEDECOD`))
- pagenization: there are many AE terms that can not be covered in one page. 
Column headers and SOC information needs to be repeated in every page.

In the code below, we count number of subjects in each AE term by SOC and treatment group.

```{r}
ana <- adae %>% 
  mutate(AESOC = toTitleCase(tolower(AESOC)), 
         AEDECOD =  toTitleCase(tolower(AEDECOD)))

t1 <-ana %>% 
  group_by(TRTAN, AESOC) %>%
  summarise(n = fmt_num(n_distinct(USUBJID), digits = 0)) %>% 
  mutate(AEDECOD = AESOC, order = 0)

t1 %>% head(4)
```
In the code below, we count number of subjects in each AE term by SOC, AE term, and treatment group.

```{r}
t2 <- ana %>% 
  group_by(TRTAN, AESOC, AEDECOD) %>%
  summarise(n = fmt_num(n_distinct(USUBJID), digits = 0)) %>%
  mutate(order = 1)

t2 %>% head(4)
```
We prepare reporting data for AE information.

```{r}
t_ae <- bind_rows(t1, t2) %>% 
        pivot_wider(id_cols = c(AESOC, order, AEDECOD), 
                    names_from = TRTAN, 
                    names_prefix = "n_",
                    values_from = n, 
                    values_fill = fmt_num(0, digits = 0)) %>%
        arrange(AESOC, order, AEDECOD) %>% 
        select(AESOC, AEDECOD, starts_with("n"))

t_ae %>% head(4)
```
We prepare reporting data for analysis population.

```{r}
t_pop <- adsl %>%
  filter(SAFFL == "Y") %>%
  count_by("TRT01AN", "SAFFL",
    var_label = "Participants in population"
  ) %>%
  mutate(
    AESOC = "pop", 
    AEDECOD = var_label) %>%
  select(AESOC,AEDECOD, starts_with("n_"))

t_pop
```
The final report data is saved in `tbl_ae_spec`.
We also add a blank row between population and AE information in the reporting table.

```{r}
tbl_ae_spec <- bind_rows(t_pop,
                         data.frame(AESOC = "pop"), 
                         t_ae) %>% 
  mutate(AEDECOD = ifelse(AEDECOD == AESOC , 
                          AEDECOD, paste0("  ", AEDECOD)))

tbl_ae_spec %>% head(4)
```
We start to define the format of the output.

To obtain the nested layout, we use `page_by` argument in `rtf_body` function.
By defining `page_by="AESOC"`, 
`r2rtf` recognize the variable as a group indicator.

After setting `pageby_row = "first_row"`, the first row is displayed as group header. If a group of information is breaked into multiple page, 
the group header row is repeated in each page by default.

We can also customize the text format by providing a matrix that have 
same dimention of the input dataset (i.e. `tbl_ae_spec`). 
In the code below, we illustrate how to display **bold** text for group headers
to highlight the nested structure of the table layout.

```{r}
n_row <- nrow(tbl_ae_spec)
n_col <- ncol(tbl_ae_spec)
id <- tbl_ae_spec$AESOC == tbl_ae_spec$AEDECOD
id <- ifelse(is.na(id), FALSE, id)

text_format   <- ifelse(id, "b", "")
```

More discussion on `page_by`, `group_by` and `subline_by` 
features can be found in the 
[`r2rtf` package website](https://merck.github.io/r2rtf/articles/example-sublineby-pageby-groupby.html.

```{r}
tbl_ae_spec %>% 
  rtf_title("Analysis of Participants With Specific Adverse Events", 
            "(Safty Analysis Population)") %>%

    rtf_colheader(" | Placebo | Xanomeline Low Dose| Xanomeline High Dose",
    col_rel_width = c(3, rep(1, 3))
  ) %>%
  rtf_colheader(" | n |  n | n ",
    border_top = "",
    border_bottom = "single",
    col_rel_width = c(3, rep(1, 3))
  ) %>%

  rtf_body(
    col_rel_width = c(1, 3, rep(1, 3)),
    text_justification = c("l", "l", rep("c", 3)),
    text_format = matrix(text_format, nrow = n_row, ncol = n_col), 
    page_by = "AESOC", 
    pageby_row = "first_row") %>%

  rtf_footnote("Every subject is counted a single time for each applicable row and column.") %>%

  rtf_encode() %>%

  write_rtf("tlf/tlf_spec_ae.rtf")

```

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("tlf/tlf_spec_ae.pdf")
```
